Coherent optical creation of a single molecule

### 2020-12-07 paper.tex
Update on Overleaf.

Here, we demonstrate coherent association of an atom pair to a weakly bound molecule
by two-photon optical Raman transfer via an electronic excited state,
as shown in Fig.~\ref{f-theory}a.
The technique does not rely on a Feshbach resonance,
molecular states bound by a only few~MHz, or a narrow excited state.
The resulting single molecule is in a well-defined internal quantum state
and predominantly in its motional ground state.
A vibrational state of the electronic excited state $\mathrm{c^3\Sigma^+}(\Omega = 1)$
serves as the intermediate state in the Raman technique,
and is chosen to minimize photon scattering during Raman Rabi oscillations.
To reduce photon scattering and sensitivity to laser intensity noise further,
we choose the initial and final states to balance the two Rabi frequencies as much as possible.
This approach applies to a variety of molecules that can be created atom-by-atom with full quantum state control.

@@

The optical Raman transfer is illustrated
by the idealized three-level system shown in Fig.~\ref{f-theory}a,
where the initial atomic state and the target weakly bound molecular state are coupled to an intermediate state by two lasers with Rabi frequencies $\Omega_a$ and $\Omega_m$ with one-photon detuning $ \Delta $, and all Rabi frequencies are population oscillation frequencies.  %(Ultimately, we calculate with contribution of all vibrational states of the excited electronic potential, $\mathrm{c^3\Sigma^+}$)
The transfer Raman Rabi frequency is given by $\Omega_a\Omega_m / (2\Delta)$~\cite{Wineland2003}.
% , is accompanied by a photon scattering rate $\Gamma_e (\Omega_a^2 + \Omega_m^2)/4\Delta^2$
% $\Gamma_e \Omega^2 / 4\Delta^2 $, where $ \Gamma_e $ is the excited state linewidth
Unlike Raman transitions in atoms, the two Rabi frequencies are greatly imbalanced ($\Omega_a/\Omega_m \ll 1$) %. Since $ \Omega_m $ is generally much larger than $ \Omega_a $
due to the small wavefunction overlap between the atomic state and the intermediate state, %the atomic state and the excited molecular state,
and scattering losses are dominated by the final state. Furthermore, the energy difference between the atomic state and target molecular state is small ($ < 1~\mathrm{GHz} $) compared to the single-photon detuning $ \Delta $ of $80$ to $200$~GHz, causing the target molecular state to scatter both beams nearly equally with a total rate $ \Gamma_e \Omega_m^2 / (2\Delta^2)$, where $ \Gamma_e $ is the excited-state linewidth~\footnote{The two beams have equal power to maximize the Raman Rabi rate at fixed total power, resulting in a factor of 2 for scattering two beams.}.
In this idealized treatment, the ratio between the Raman Rabi frequency and the scattering rate is $ \Omega_a/\Omega_m \times \Delta/\Gamma_e $, which limits the transfer efficiency into the molecular state. At the same time, the intensity-stability requirement is determined by the ratio of Raman Rabi frequency to light shift ($\Omega_a/\Omega_m$)~\footnote{There is an additional factor of 2, with both beams at equal power,
  to account for the Stark shift caused by both beams.}. Notably, both figures of merit improve with a larger ratio $\Omega_a/\Omega_m$. %, depends on the ratio of the two matrix elements and how far detuned the laser is from the transition in units of the linewidth.
% However, the detuning cannot be too large, since that will reduce the Raman Rabi frequency.

Earlier experiments used weakly bound excited states as the intermediate state
in the Raman transition to ensure a large Raman Rabi frequency~\cite{Wynar2000,Rom2004}.
However, a complete picture includes both the many vibrational levels
of the excited electronic state and the atomic continuum.
The total scattering rate and Raman Rabi rate become a sum of the scattering rates
and Raman Rabi rates over all possible intermediate states.
As there is large overlap between the target molecular state and weakly bound excited states, intermediate states that are closer to the dissociation threshold result in a large scattering rate.
% in large incoherence and loss to other molecular states.
This scattering is approximately proportional to $1/\delta_{\mathrm{thresh}}^2$,
where $\delta_{\mathrm{thresh}}$ is the detuning from the dissociation threshold,
which is smaller for intermediate states that are deeply bound \footnote{See Supplementary Material}. \TR{Cite supplement or maybe Fig. S1 can go into Fig. 1b since it shows $\Gamma_s\sim \delta_{thresh}^{-2}$ nicely.}.

@@

\begin{figure}[t!]
  \includegraphics[width=0.48\textwidth]{imgs/fig-raman.pdf}
  \caption{Coherent transfer of atoms to molecules.  The molecular state is dark to the imaging step and corresponds to zero signal.
    (a) Raman difference frequency scans for various durations showing the resonance as a frequency offset from $770.2$~MHz, near the calculated value of $763$~MHz.
    (b) Raman pulse-length scan on resonance.
    A decaying Rabi oscillation shows the coherence of
    the Raman transfer process.
    A model is fitted to (a) and (b) to determine
    the Raman Rabi frequency and loss rates.
    \emph{Inset:} Geometry and polarization of trap and Raman beam relative to the magnetic field.
    The X.yy~mW beam is focused to a waist of $0.9~\mathrm{\mu m}$
    that confines the atoms and molecule.
    A perpendicular $B_0=8.83(2)~\mathrm{G}$ magnetic field
    defines the quantization axis and the atoms experience predominantly $\pi$-polarized light.
    \label{f-raman}}
\end{figure}
%% Transfer scheme
% After the atoms are prepared in the $\ket{\uparrow_{\Na}\uparrow_{\Cs}} $ hyperfine combination, we then perform the Raman transfer.
To transfer the atom pair into the target weakly bound molecular state,
we modulate the tweezer beam with a second frequency near $770$~MHz, as shown in Fig.~\ref{f-beampath}.
The dual use of the tweezer beam for confinement and Raman transfer not only minimizes photon scattering,
but also allows a tight focus to minimize the transfer duration. A tweezer frequency far detuned (by $-151~\mathrm{GHz}$) from $v' = 0 $ ($<50$~MHz natural linewidth) reduces resonant scattering~\cite{Liu2019}.
% Furthermore, two filters, each with a linewidth (FWHM) of $50~\mathrm{GHz}$,
% clean the laser spectrum and prevent broadband noise from causing unwanted excitation.
% As shown in Fig.~\ref{f-beampath}, one filter immediately follows the laser, while the second filter precedes the microscope objective for final cleanup of the laser spectrum.
% The filters reduce photon scattering rate by a factor of $2$ (\textcolor{red}{KW: Since we mention this again later when we explore the causes of our lower than expected transfer efficiency, do we need to mention this reduction here? or the possible causes?}). While we have not fully characterized the sources
% of broadband noise, possibilities include amplified spontaneous emission (ASE) from the laser and fiber nonlinearities.
% After the total tweezer power is set to the desired value,
% we smoothly ramp down the power of one frequency in the tweezer
% while simultaneously ramping up the power of the other frequency
% so that the total tweezer power remains unchanged.
% Both frequencies are kept on for a specified duration before the process is reversed
% and the tweezer returns to a single frequency.~\TR{???}

>>>>

Here, we demonstrate coherent association of an atom pair to a weakly bound molecule
by two-photon optical Raman transfer via an electronic excited state,
as shown in Fig.~\ref{f-theory}a.
The technique does not rely on a Feshbach resonance,
molecular states bound by a only few~MHz, or a narrow excited state.
The resulting single molecule is in a well-defined internal quantum state
and predominantly in its motional ground state.
A vibrational state of the electronic excited state $\mathrm{c^3\Sigma^+}(\Omega = 1)$
serves as the intermediate state in the Raman technique,
and is chosen to minimize photon scattering during Raman Rabi oscillations.
To reduce photon scattering and sensitivity to laser intensity noise further,
we choose the initial and final states that balance the two Rabi frequencies as much as possible.
This approach applies to a variety of molecules that can be created atom-by-atom with full quantum state control.

@@

The optical Raman transfer is illustrated
by the idealized three-level system shown in Fig.~\ref{f-theory}a,
where the initial atomic state and the target weakly bound molecular state are coupled to an intermediate state by two lasers with Rabi frequencies $\Omega_a$ and $\Omega_m$, one-photon detuning $ \Delta $, and all Rabi frequencies are population oscillation frequencies.  %(Ultimately, we calculate with contribution of all vibrational states of the excited electronic potential, $\mathrm{c^3\Sigma^+}$)
The transfer Raman Rabi frequency is given by $\Omega_a\Omega_m / (2\Delta)$~\cite{Wineland2003}.
% , is accompanied by a photon scattering rate $\Gamma_e (\Omega_a^2 + \Omega_m^2)/4\Delta^2$
% $\Gamma_e \Omega^2 / 4\Delta^2 $, where $ \Gamma_e $ is the excited state linewidth
Unlike Raman transitions in atoms, the one-photon Rabi frequencies are greatly imbalanced ($\Omega_a/\Omega_m \ll 1$) %. Since $ \Omega_m $ is generally much larger than $ \Omega_a $
due to the small wavefunction overlap between the atomic state and the intermediate state, %the atomic state and the excited molecular state,
and scattering losses are dominated by the final state. Furthermore, the energy difference between the atomic state and target molecular state is small ($ < 1~\mathrm{GHz} $) compared to the single-photon detuning of $80$ to $200$~GHz, causing the target molecular state to scatter both beams nearly equally with a total rate $ \Gamma_e \Omega_m^2 / (2\Delta^2)$, where $ \Gamma_e $ is the excited-state linewidth~\footnote{The two beams have equal power to maximize the Raman Rabi rate at fixed total power, resulting in a factor of 2 for scattering two beams.}.
In this idealized treatment, the ratio between the Raman Rabi frequency and the scattering rate is $ \Omega_a/\Omega_m \times \Delta/\Gamma_e $, which limits the transfer efficiency into the molecular state. At the same time, the intensity-stability requirement is determined by the ratio of Raman Rabi frequency to light shift ($\Omega_a/\Omega_m$)~\footnote{There is an additional factor of 2, with both beams at equal power,
  to account for the Stark shift caused by both beams.}. Notably, both figures of merit improve with a larger ratio $\Omega_a/\Omega_m$. %, depends on the ratio of the two matrix elements and how far detuned the laser is from the transition in units of the linewidth.
% However, the detuning cannot be too large, since that will reduce the Raman Rabi frequency.

Earlier experiments used weakly bound excited states as the intermediate state
in the Raman transition to ensure a large Raman Rabi frequency~\cite{Wynar2000,Rom2004}.
However, a complete picture includes both the many vibrational levels
of the excited electronic state and the atomic continuum.
Then the scattering and Raman Rabi rates are sums over all possible intermediate states.
As there is large overlap between the target molecular state and weakly bound excited states, intermediate states that are closer to the dissociation threshold result in a large scattering rate.
% in large incoherence and loss to other molecular states.
This scattering is approximately proportional to $1/\delta_{\mathrm{thresh}}^2$,
where $\delta_{\mathrm{thresh}}$ is the detuning from the dissociation threshold,
which is smaller for intermediate states that are deeply bound \footnote{See Supplementary Material}. \TR{Cite supplement or maybe Fig. S1 can go into Fig. 1b since it shows $\Gamma_s\sim \delta_{thresh}^{-2}$ nicely.}.

@@

\begin{figure}[t!]
  \includegraphics[width=0.48\textwidth]{imgs/fig-raman.pdf}
  \caption{Coherent transfer of atoms to molecules.  The molecular state is dark to the imaging step and corresponds to zero signal.
    (a) Raman difference frequency scans for various durations showing the resonance as a frequency offset from $770.2$~MHz, near the calculated value of $763$~MHz.
    (b) Raman pulse-length scan on resonance.
    A decaying Rabi oscillation shows the coherence of
    the Raman transfer process.
    A model is fitted to (a) and (b) to determine
    the Raman Rabi frequency and loss rates.
    \emph{Inset:} Geometry and polarization of trap and Raman beam relative to the magnetic field.
    The X.yy~mW beam is focused to a waist of $0.9~\mathrm{\mu m}$
    that confines the atoms and molecule.
    A perpendicular $B_0=8.83(2)~\mathrm{G}$ magnetic field
    defines the quantization axis and the atoms experience predominantly $\pi$-polarized light.
    \label{f-raman}}
\end{figure}
%% Transfer scheme
% After the atoms are prepared in the $\ket{\uparrow_{\Na}\uparrow_{\Cs}} $ hyperfine combination, we then perform the Raman transfer.
To transfer the atom pair into the target weakly bound molecular state,
we modulate the tweezer beam with a second frequency near $770$~MHz, as shown in Fig.~\ref{f-beampath}.
The dual use of the tweezer beam for confinement and Raman transfer not only minimizes photon scattering,
but also allows a tight focus to minimize the transfer duration. A tweezer frequency far detuned (by $-151~\mathrm{GHz}$) from $v' = 0 $ ($<50$~MHz natural linewidth) reduces resonant scattering~\cite{Liu2019}.
Furthermore, two filters, each with a linewidth (FWHM) of $50~\mathrm{GHz}$,
 clean the laser spectrum and prevent broadband noise from causing unwanted excitation.
 As shown in Fig.~\ref{f-beampath}, one filter immediately follows the laser, while the second filter precedes the focusing objective for final cleanup of the laser spectrum.
% The filters reduce photon scattering rate by a factor of $2$ (\textcolor{red}{KW: Since we mention this again later when we explore the causes of our lower than expected transfer efficiency, do we need to mention this reduction here? or the possible causes?}). While we have not fully characterized the sources
% of broadband noise, possibilities include amplified spontaneous emission (ASE) from the laser and fiber nonlinearities.
% After the total tweezer power is set to the desired value,
% we smoothly ramp down the power of one frequency in the tweezer
% while simultaneously ramping up the power of the other frequency
% so that the total tweezer power remains unchanged.
% Both frequencies are kept on for a specified duration before the process is reversed
% and the tweezer returns to a single frequency.~\TR{???}
